---
title: "Recalibration of datasets"
description: >
  Learn how to get recalibrate datasets using TrackMateR.
output:
  rmarkdown::html_vignette:
    fig_width: 8
    fig_height: 8
vignette: >
  %\VignetteIndexEntry{Recalibration of datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In the original example, we saw how to load one dataset and recalibrate.

```{r basic-example, eval=FALSE}
library(ggplot2)
library(TrackMateR)
# an example file is provided, otherwise use file.choose()
xmlPath <- system.file("extdata", "ExampleTrackMateData.xml", package="TrackMateR")
# read the TrackMate XML file into R using
tmObj <- readTrackMateXML(XMLpath = xmlPath)
# Pixel size is 0.04 um and original data was 1 pixel, xyscalar = 0.04
tmObj <- correctTrackMateData(dataList = tmObj, xyscalar = 0.04, xyunit = "um")
```

For the example file, the spatial units are 1 pixel and not the real value which is 0.04 um per pixel.
So to convert we can specify `xyscalar` to change 1 to 0.04 and specify `xyunit` to change from "pixel" to "um".
If the file had 2 pixels as the spatial unit (for some reason), to convert to 0.04 um we would need to use `xyscalar = 0.02, xyunit = "um"`

The units (`xyunit` and `tunit`) can be any arbitrary string.
The use of "um" is preferred over "µm" because the mu character prints inconsistently in R graphics across platforms.

If the time scaling of the file is 0.07 s and we would prefer ms as the unit,

```{r eval=FALSE}
tmObj <- correctTrackMateData(dataList = tmObj, xyscalar = 0.04, xyunit = "um", tscalar = 1000, tunit = "ms")
```

this is because the file is loaded in and scaled at 1 s, so the conversion needs to be 1000 ms.
If we wanted to switch to minutes, we would use `tscale = 1/60, tunit = "min"`.

If the time scaling is incorrect, e.g. 1 frame, we would need to rescale using `tscale = 0.07, tunit = "s"`, `tscale = 70, tunit = "ms"`, or `tscale = 0.07/60, tunit = "min"`.


## Recalibrating TrackMate files

When using `compareDatasets()`, it is possible to recalibrate TrackMate XML files _en masse_ by using a CSV file in the condition subfolder containing the XML files, see `vignette("comparison")`.
To recalibrate, a separate CSV file is needed for each condition.
Any conditions that do not need recalibrating simply do not need a CSV file present.
The CSV file can have any name but must follow this format.

```{text}
value,unit
0.04,um
0.07,s

```

Using this file, all XML files in the folder will be recalibrated if they are not already at 0.04 micron pixel size and 0.07 s per frame.
Adjustments smaller than 2.5% are ignored.

**Note that the calibration file should give the correct final scaling required for all files in the folder.**

When the rescaling is done using a file like this, TrackMateR will work out the scalar for you.
So the file needs to specify what final scaling is required.

In the situation where you have datasets with differing (but correct) timescales and spatial units need to be corrected, use 0 for the value corresponding to time (time rescaling will be ignored) and spatial rescaling will go ahead.
Similarly, if you have differening (but correct) spatial scaling between datasets you can rescale time only by using 0 for the spatial value.
If you have datasets that need calibrating differently, you will need to group them into condition-level subfolders with a single csv file to specify the correct parameters.
Alternatively, consider going back and retracking in TrackMate and specifying the correct parameters.

## Limitations

The recalibration is not smart, it doesn't recognise the units and calculate based off that.
For example, it doesn't know that 1000 ms are 1 s.
So if you use the recalibration file method, be aware of this limitation.
