---
title: "Introduction to TrackMateR"
description: >
  Learn how to get started with the basics of TrackMateR.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to TrackMateR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  message = FALSE
)
```


TrackMateR is an R package to analyse TrackMate XML outputs.

[TrackMate](https://imagej.net/plugins/trackmate/) is a single-particle tracking plugin for ImageJ/Fiji.
The standard output from a tracking session is in TrackMate XML format.

The goal of this R package is to import all of the data associated with the final filtered tracks in TrackMate for further analysis and visualization in R.

## Installation

Once you have installed [R](https://cran.rstudio.com) and [RStudio Desktop](https://www.rstudio.com/products/rstudio/download/), you can install TrackMateR using devtools

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("quantixed/TrackMateR")
```

## An Example

A basic example is to load one TrackMate XML file and analyse it.
This would typically be a tracking session associated with a single movie (referred to here as a single dataset).

```{r example}
library(ggplot2)
library(TrackMateR)
# an example file is provided, otherwise use file.choose()
xmlPath <- system.file("extdata", "ExampleTrackMateData.xml", package="TrackMateR")
# read the TrackMate XML file into R using
tmObj <- readTrackMateXML(XMLpath = xmlPath)
# perhaps the data we loaded in was not scaled properly
# Pixel size is 0.04 um and original data was 1 pixel, xyscalar = 0.04
tmObj <- correctTrackMateData(dataList = tmObj, xyscalar = 0.04, xyunit = "um")
# now we can have a look at the tracks
plot_tm_allTracks(tmObj)
```

After reading in the TrackMate file with `readTrackMateXML()` we have an object which we can visualise using `plot_tm_allTracks()`.

TrackMateR can generate several different types of plot individually using commands like `plot_tm_allTracks()` or it can make them all automatically and create a report for you.

```{r, message = FALSE}
reportDataset(tmObj)
```

## A more advanced example

Perhaps the default options do not suit your needs.
With just a few more lines of code, you can tweak the report to do the analysis you want.

Below you will see that the dataset we read in was not scaled correctly.
After reading in, we rescale it and then we start to generate the report.

The report shows analysis of mean squared displacement, jump distance and track density, among other plots.
These three things must be calculated first using `calculateMSD()`, `calculateJD()` and `calculateTrackDensity`.
This allows you to have some control over the analysis, for example setting the search radius for Track Density.
Finally, these items can be fed into `makeSummaryReport()` as shown below, and the report is created.


```{r, message = FALSE}
# we can get a data frame of the correct TrackMate data 
tmDF <- tmObj[[1]]
# and a data frame of calibration information
calibrationDF <- tmObj[[2]]
# we can calculate mean squared displacement
msdObj <- calculateMSD(df = tmDF, method = "ensemble", N = 3, short = 8)
# and jump distance
jdObj <- calculateJD(dataList = tmObj, deltaT = 1)
# and look at the density of tracks
tdDF <- calculateTrackDensity(dataList = tmObj, radius = 1.5)
# if we extract the name of the file
fileName <- tools::file_path_sans_ext(basename(xmlPath))
# we can send all these things to makeSummaryReport() to get a nice report of our dataset
makeSummaryReport(tmList = tmObj, msdList = msdObj, jumpList = jdObj, tddf = tdDF,
titleStr = "Report", subStr = fileName, auto = FALSE)
```

This vignette described the analysis of a single dataset.
For analysis of more than one dataset, see `vignette("comparison")`
